---
title: "PeerAssessment"
author: "roberto avendano"
date: "Sunday, August 17, 2014"
output: html_document
---
## Loading and preprocessing the data :

Reading the data from the file activity.csv in the current directory:
```{r}
getwd()
```

```{r}
table<- read.csv("activity.csv")
table$date <- as.Date(table$date, "%Y-%m-%d")
## table2<-table[!is.na(table$steps),]

```

## What is mean total number of steps taken per day?

```{r}
StepsPerDay <- aggregate(steps~date, table, sum)
hist(StepsPerDay$steps, xlab="steps", main="Steps per Day", breaks=20)
mean(StepsPerDay$steps)
median(StepsPerDay$steps)
```

## What is the average daily activity pattern?

```{r}
intervals <- aggregate(steps~interval, table, mean)
colnames(intervals)[2] <- "mean_steps"
plot(intervals, type="l", ylab="Average Steps", main="Average steps per interval")
```
Thus the following interval  contains, on average, the greatest number of steps.

```{r}
max.interval <- intervals[intervals$mean.steps == max(intervals$mean.steps), 1]
intervals[intervals$interval == max.interval, ]
```

`r max.interval`

##  Imputing missing values

The number of missing values is :

```{r}
 sum(is.na(table$steps))
```

Let´s replace the NAs with the mean value of its corresponding dayss.

```{r}
table2 <- merge(table, intervals, by="interval")
table2 <- table2[order(table2$date), ]
table2$steps[is.na(table2$steps)] <- table2$mean_steps[is.na(table2$steps)]
activity2$mean.steps <- NULL
```

```{r}
StepsPerDay2 <- aggregate(steps~date, table2, sum)
hist(StepsPerDay2$steps, xlab="steps", main="Steps per Day (modified data)", breaks=20)
```


```{r}
mean(StepsPerDay2$steps)
```

```{r}
median(StepsPerDay2$steps)
```
 
 
The mean and median values are very similar, but the tallest bar of the histogram has been lengthened. The reason for this is that by filling in the NA values added more days with the average number of daily steps.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
table2$day <- weekdays(table2$date)
table2$daytype <- "weekday"
table2$daytype[table2$day == "Saturday"] <- "weekend"
table2$daytype[table2$day == "Sunday"]   <- "weekend"
table2$day <- NULL
table2$daytype <- factor(activity2$daytype)

intervals.weekday <- aggregate(steps~interval, table2[table22$daytype == "weekday", ], mean)
intervals.weekend <- aggregate(steps~interval, table2[table2$daytype == "weekend", ], mean)
intervals.weekday$daytype <- "weekday"
intervals.weekend$daytype <- "weekend"
intervals.both <- rbind(intervals.weekday, intervals.weekend)
intervals.both$daytype <- factor(intervals.both$daytype)
library(lattice)
xyplot(steps~interval | daytype, data=intervals.both, type="l", layout=c(1,2))
```
 
 
 
